GA-based Training-Free NAS Algorithm with Hybrid Score Function

### 2023-03-16 paper/main.tex
update related work (figure of ni && advices)

    Most neural architecture searches (NASs) are time-consuming caused by 
    the fact that, during the searching, a candidate architecture must be 
    trained to evaluate how good of this architecture. This is why some of 
    training-free NAS algorithms have been proposed in recent years.

    Although the training-free NASs are typically faster than training-based 
    NAS method, however, the correlation between score value and the result 
    of an architecture is not well enough in most cases.

    To address this problem, we propose a genetic-based training-free NAS 
    algorithm with hybrid training-free score function, which combines three 
    highly heterogeneous training-free score functions to evaluate an architecture. 
    In this method, the genetic algorithm plays a role to guide the searches 
    of NAS algorithm while the hybrid training-free score function plays the 
    role to evaluate a new candidate architecture during the convergencr process 
    of GA. More precisely, the first score function is noise immunity for 
    neural architecture search without search (NINASWOT), as an evaluation of 
    pattern recognition ability, second one is maximum-entropy detection (MAE-DET), 
    as an evaluation of the entropy of an architecture and the third one is 
    condition number of neural tangent kernel (NTK), as an evaluation of 
    the speed of converge.

    To evaluate the performance of the proposed algorithm, we compared it 
    with several NAS algorithms, including weight-sharing methods, 
    non-weight-sharing methods, and neural architecture search without training (NASWOT). 
    We expect develope a faster and more accurate training-free NAS algorithm.

@@

% introduction
    Neural architecture search (NAS) has recently drawn a big amount of 
    attention, since the ability to automatically design a "good" neural 
    architecture. By leveraging machine learning algorithms \cite{https://doi.org/10.48550/arxiv.1611.01578}, 
    NAS algorithms can explore a search space, which is comprised of numerous 
    potential architectures, to find out a good architectures that outperform 
    those designed by human experts. In recent years, the use of NAS is 
    widespread, from object detection \cite{https://doi.org/10.48550/arxiv.2111.13336}, 
    image recognition \cite{https://doi.org/10.48550/arxiv.2006.04647} 
    and speech recognition \cite{https://doi.org/10.48550/arxiv.2011.05649} 
    \cite{mehrotra2021nasbenchasr} to natural language 
    processing. \cite{jiang-etal-2019-improved} 
    \cite{https://doi.org/10.48550/arxiv.2006.07116} 
    \cite{https://doi.org/10.48550/arxiv.2005.14187}

    Despite the promising results of NAS, there are still many challenges 
    to conquer. One major problem is the extremely high computational 
    cost to search for an optimal architecture, which can make NAS impractical 
    for real-world applications, particularly on resource-constrained 
    platforms like embedded system. The reason why NAS is costly is that 
    during the searching, a candidate architecture must be trained to 
    evaluate how good of this architecture.

    To overcome this challenge, recent works developed and proposed lots 
    of method which is so called training-free NAS. 
    % NINASWOT
    For example, Mellor et al. \cite{https://doi.org/10.48550/arxiv.2006.04647} proposed the 
    measurement of the correlation between the binary activation paterns, 
    induced by the untrained network at two inputs, named as neural architecture 
    search without training (NASWOT). The correlation is defined as the 
    logarithmic determinant of the kernel matrix, as follows:
    \begin{equation}
        s=\log\lvert K_H\rvert
    \end{equation}
    where $K_H$ is the kernel matrix.
    Later, Wu et al. \cite{10.1145/3491396.3506510} found a high score 
    obtained by such a function may not correspond to a high-performence 
    model. Thus, they additionally applied noisy immunity method on NASWOT, 
    named as noisy immunity for NASWOT (NINASWOT).
    % NTK
    For another example, Chen et al. proposed to compute the condition 
    number of neural tangent kernel (NTK) \cite{https://doi.org/10.48550/arxiv.2102.11535} 
    \cite{https://doi.org/10.48550/arxiv.2203.09137} \cite{https://doi.org/10.48550/arxiv.2109.00817}, 
    which is defined as 
    \begin{equation}
        \mathcal{K_N}=\frac{\lambda_{\textrm{max}}(\hat\Theta)}{\lambda_{\textrm{min}}(\hat\Theta)}
    \end{equation}
    and used as the score to estimate the trainability of an architecture.
	% unknown

@@

    % naswot
    In \cite{https://doi.org/10.48550/arxiv.2006.04647}, Mellor et al. proposed a 
    score function without the requirement for training. Here we call it neural 
    architecture search without training (NASWOT) score function. 
    Consider a mini-batch of data $X=\{x_i\}^N_{i=1}$ passing through a neural network 
    architecture as $f(x_i)$. The activated ReLU units in every layer of the architecture 
    form a binary code $c_i$ that define the linear region.
    The correlation between binary codes for the whole mini-batch can be examined 
    by computing the kernel matrix
    \begin{equation}
        K_H=\begin{pmatrix}N_A-d_H(c_1,c_2)&\cdots&N_A-d_H(c_1,c_N)\\\vdots&\ddots&\vdots\\N_A-d_H(c_N,c_1)&\cdots&N_A-d_H(c_N,c_1)\end{pmatrix}
    \end{equation}
    where $N_A$ is the number of ReLU units and $d_H(c_i,c_j)$ is the hamming 
    distance between the binary code $c_i$ and $c_j$. 
    With the kernel matrix, the score of an architecture can be evaluate as 
    follow: 
    \begin{equation}
        s=\log\lvert K_H\rvert
    \end{equation}
    The rationale behind is to see how dissimilar the linear region activated by 
    the ReLU units between two inputs. An architecture shall learn better 
    when inputs are well separated.

    % ninaswot
    Based on the work of Mellor et al., Wu et al. \cite{10.1145/3491396.3506510} 
    found, in some case, an architecture with high score may put different 
    input data, which are originally in the same class, to different classes. 
    To fix this defect, Wu et al. proposed using noise immunity to additionally 
    evaluate an architecture. The score function picks a mini-batch of data, 
    denoted $X$, and then applies Gaussion noise on it, denoted $X'$ and defined 
    by $X'=X+z$ where $z$ is the Gaussion noise. By passing $X$ and $X'$ through 
    the untrained architecture, the sum of the square differences between outputs 
    $O=\begin{matrix}o_1,o_2,\cdots,o_C\end{matrix}$ and 
    $O'=\begin{matrix}o_1',o_2',\cdots,o_C'\end{matrix}$ can be calculated as 
    follows:
    \begin{equation}
        n=-\sum^C_{n=1}(o_i-o_i')^2
    \end{equation}
    where $C$ is the number of classes determined by the input data.
    According to the experiment outcome of Wu et al., noise immunity can separate 
    those are unable to distinguish $X$ and $X'$ from architectures with high NASWOT 
    score, which just complements the NASWOT score funcion.

    % ntk
    On the other hand, instead of using the correlation of binary activation 
    patterns as score. Chen et al. proposed using the condition number of NTK 
    as an evaluation of the trainability, defined as
    \begin{equation}
        \mathcal{K_N}=\frac{\lambda_{\textrm{max}}(\hat\Theta)}{\lambda_{\textrm{min}}(\hat\Theta)}
    \end{equation}
    where $\lambda_{\textrm{max}}(\hat\Theta)$ and $\lambda_{\textrm{min}}(\hat\Theta)$ 
    are the maximum and minimum eigenvalues of NTK ($\hat\Theta$) respectively.
    The rationale is based on the work of Lee et al. \cite{Lee_2020} and 
    Xiao et al. \cite{https://doi.org/10.48550/arxiv.1912.13053}, which concluded briefly here, 
    the training dynamics of a wide neural network can be written in terms of the 
    spectrum of the NTK:
    \begin{equation}\label{eq:6}
        \mu_t(\textbf{X}_{\textrm{train}})_i=(\textbf{I}-e^{-\eta\lambda_it})\textbf{Y}_{\textrm{train,}i}
    \end{equation}
    where $\lambda_i$ are the eigenvalues of $\hat\Theta_{\textrm{train}}$ 
    and they're ordered as $\lambda_0\geq\cdots\geq\lambda_m$. 
    According to the hypothesis proposed by Lee et al. \cite{Lee_2020}, the maximum 
    feasible learning rate scale as $\eta\sim2/\lambda_0$. Plug this scaling for 
    $\eta$ in Eq. \ref{eq:6}, we see that the $\lambda_m$ will converge exponentially 
    at a rate given by $1/K_N$ where $K_N=\frac{\lambda_0}{\lambda_m}$.
    Then it can be concluded that if $K_N$ is lower, the corresponding architecture is 
    more trainable.

>>>>

    % dont we
    % To address
    % To evaluate
    Although the training-free NASs are typically faster than training-based 
    NAS method, however, the correlation between score value and the result 
    of an architecture is not well enough in most cases.
    To address this problem, we propose a genetic-based training-free NAS 
    algorithm with hybrid training-free score function, which combines three 
    highly heterogeneous training-free score functions to evaluate an architecture. 
    In this method, the genetic algorithm plays a role to guide the searches 
    of NAS algorithm while the hybrid training-free score function plays the 
    role to evaluate a new candidate architecture during the convergencr process 
    of GA. More precisely, the first score function is noise immunity for 
    neural architecture search without search (NINASWOT), as an evaluation of 
    pattern recognition ability, second one is condition number of neural 
    tangent kernel (NTK), as an evaluation of trainability.

    To evaluate the performance of the proposed algorithm, this paper compared 
    it with several NAS algorithms, including weight-sharing methods, 
    non-weight-sharing methods, and neural architecture search without training (NASWOT). 
    We expect develope a faster and more accurate training-free NAS algorithm.

@@

% introduction
    Neural architecture search (NAS) has recently drawn a big amount of 
    attention, since the ability to automatically design a "good" neural 
    architecture. By leveraging machine learning algorithms \cite{https://doi.org/10.48550/arxiv.1611.01578}, 
    NAS algorithms can explore a search space, which is comprised of numerous 
    potential architectures, to find out a good architectures that outperform 
    those designed by human experts. "In recent years", the use of NAS is 
    widespread, from object detection \cite{https://doi.org/10.48550/arxiv.2111.13336}, 
    image recognition \cite{https://doi.org/10.48550/arxiv.2006.04647} 
    and speech recognition \cite{https://doi.org/10.48550/arxiv.2011.05649} 
    \cite{mehrotra2021nasbenchasr} to natural language 
    processing. \cite{jiang-etal-2019-improved} 
    \cite{https://doi.org/10.48550/arxiv.2006.07116} 
    \cite{https://doi.org/10.48550/arxiv.2005.14187}

    % reduce 
    Despite the promising results of NAS, there are still many challenges 
    to conquer. One major problem is the extremely high computational 
    cost to search for an optimal architecture, which can make NAS impractical 
    for real-world applications, particularly on resource-constrained 
    platforms like embedded system. The reason why NAS is costly is that 
    during the searching, a candidate architecture must be trained to 
    evaluate how good of this architecture.

    To overcome this challenge, recent works developed and proposed lots 
    of method which is so called training-free NAS. 
    % NINASWOT
    For example, Mellor et al. \cite{https://doi.org/10.48550/arxiv.2006.04647} proposed the 
    measurement of the correlation between the binary activation paterns, 
    induced by the untrained network at two inputs, named as neural architecture 
    search without training (NASWOT). 
    Later, Wu et al. \cite{10.1145/3491396.3506510} found a high score 
    obtained by such a function may not correspond to a high-performence 
    model. Thus, they additionally applied noisy immunity method on NASWOT, 
    named as noisy immunity for NASWOT (NINASWOT).
    % NTK
    For another example, Chen et al. proposed to compute the condition 
    number of neural tangent kernel (NTK) \cite{https://doi.org/10.48550/arxiv.2102.11535} 
    \cite{https://doi.org/10.48550/arxiv.2203.09137} \cite{https://doi.org/10.48550/arxiv.2109.00817}, 
    which is defined as % delete math formula
    \begin{equation}
        \mathcal{K_N}=\frac{\lambda_{\textrm{max}}(\hat\Theta)}{\lambda_{\textrm{min}}(\hat\Theta)}
    \end{equation}
    and used as the score to estimate the trainability of an architecture.
	% unknown

@@

    % naswot
    In \cite{https://doi.org/10.48550/arxiv.2006.04647}, Mellor et al. proposed a 
    score function without the requirement for training. Here we call it neural 
    architecture search without training (NASWOT) score function. 
    Consider a mini-batch of data $X=\{x_i\}^N_{i=1}$ passing through a neural network 
    architecture as $f(x_i)$. The activated ReLU units in every layer of the architecture 
    form a binary code $c_i$ that define the linear region.
    The correlation between binary codes for the whole mini-batch can be examined 
    by computing the kernel matrix % add figure
    \begin{equation}
        K_H=\begin{pmatrix}N_A-d_H(c_1,c_2)&\cdots&N_A-d_H(c_1,c_N)\\\vdots&\ddots&\vdots\\N_A-d_H(c_N,c_1)&\cdots&N_A-d_H(c_N,c_1)\end{pmatrix}
    \end{equation}
    where $N_A$ is the number of ReLU units and $d_H(c_i,c_j)$ is the hamming 
    distance between the binary code $c_i$ and $c_j$. 
    With the kernel matrix, the score of an architecture can be evaluate as 
    follow: 
    \begin{equation}
        s=\log\lvert K_H\rvert
    \end{equation}
    The rationale behind is to see how dissimilar the linear region activated by 
    the ReLU units between two inputs. An architecture shall learn better 
    when inputs are well separated.

    % ninaswot
    \begin{figure*}[htb]
        \vspace{-\baselineskip}
        \resizebox{1.0\textwidth}{!}{\includegraphics{asset/ni.pdf}}
        \caption{A simple example to illustrate the procedure of noise immunity.}
        \label{fig:ni}
        \centering
        \vspace{-\baselineskip}
    \end{figure*}
    Based on the work of Mellor et al., Wu et al. \cite{10.1145/3491396.3506510} 
    found, in some case, an architecture with high score may put different 
    input data, which are originally in the same class, to different classes. 
    To fix this defect, Wu et al. proposed using noise immunity to additionally 
    evaluate an architecture. The score function picks a mini-batch of data, 
    denoted $X$, and then applies Gaussion noise on it, denoted $X'$ and defined 
    by $X'=X+z$ where $z$ is the Gaussion noise. By passing $X$ and $X'$ through 
    the untrained architecture, the sum of the square differences between outputs 
    $O=\begin{matrix}o_1,o_2,\cdots,o_C\end{matrix}$ and 
    $O'=\begin{matrix}o_1',o_2',\cdots,o_C'\end{matrix}$ can be calculated as 
    follows:
    \begin{equation}
        n=-\sum^C_{n=1}(o_i-o_i')^2
    \end{equation}
    where $C$ is the number of classes determined by the input data.
    According to the experiment outcome of Wu et al., noise immunity can separate 
    those are unable to distinguish $X$ and $X'$ from architectures with high NASWOT 
    score, which just complements the NASWOT score funcion.

    % ntk
    On the other hand, instead of using the correlation of binary activation 
    patterns as score. Chen et al. proposed using the condition number of NTK 
    as an evaluation of the trainability, defined as % add figure
    \begin{equation}
        \mathcal{K_N}=\frac{\lambda_{\textrm{max}}(\hat\Theta)}{\lambda_{\textrm{min}}(\hat\Theta)}
    \end{equation}
    where $\lambda_{\textrm{max}}(\hat\Theta)$ and $\lambda_{\textrm{min}}(\hat\Theta)$ 
    are the maximum and minimum eigenvalues of NTK ($\hat\Theta$) respectively.
    The rationale is based on the work of Lee et al. \cite{Lee_2020} and 
    Xiao et al. \cite{https://doi.org/10.48550/arxiv.1912.13053}, which concluded briefly here, 
    the training dynamics of a wide neural network can be written in terms of the 
    spectrum of the NTK:
    \begin{equation}\label{eq:6}
        \mu_t(\textbf{X}_{\textrm{train}})_i=(\textbf{I}-e^{-\eta\lambda_it})\textbf{Y}_{\textrm{train,}i}
    \end{equation}
    where $\lambda_i$ are the eigenvalues of $\hat\Theta_{\textrm{train}}$ 
    and they're ordered as $\lambda_0\geq\cdots\geq\lambda_m$. 
    According to the hypothesis proposed by Lee et al. \cite{Lee_2020}, the maximum 
    feasible learning rate scale as $\eta\sim2/\lambda_0$. Plug this scaling for 
    $\eta$ in Eq. \ref{eq:6}, we see that the $\lambda_m$ will converge exponentially 
    at a rate given by $1/K_N$ where $K_N=\frac{\lambda_0}{\lambda_m}$.
    Then it can be concluded that if $K_N$ is lower, the corresponding architecture is 
    more trainable.
